Buildings in the City of Seattle by Nicholas Huh
========================================================


```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(XML)
library(plyr)
library(maps)
library(mapproj)
library(ggmap)
library(dplyr)
library(stringr)
library(gridExtra)
library(corrplot)

```

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
setwd('C:/Users/admin/Desktop/Nanodegree/R/Project 5')
df <- read.csv('popden_county.csv')
```

# Abstract

The city of Seattle is known as a busy city. The population has been increasing, and new buildings are being constructed and old buildings are being demolished. If you want to see if the city of Seattle is really being developed, you need to know numbers: population, number of buildings being constructed or demolished, the purpose of buildings being constructed or demolished. Using the package ggmap, we will explore the dataset of demography in the state of Washington and building information in the city of Seattle.


# Data Wrangling

```{r echo=FALSE, Data_Cleaning}
str(df)
levels(df$subregion) <- tolower(levels(df$subregion))
df$subregion <- as.character(df$subregion)
df$area <- as.character(df$area)
df$density <- as.character(df$density)

str(df)
```

# Summary of the dataset

Subregion is each subregion in the state of Washington, including the subregion of King, where the city of Seattle is located.\

Area is the total land area of each subregion, in square mile in 2010.\

Density is the population density in 2017.\

Lastly I used Pandas in python to retrieve filtered information as described above.\

You can look up the detail of dataset from the Office of Financial Management in Washington: http://www.ofm.wa.gov/pop/popden/\


#Mapping and plotting by area, population, and density

```{r echo=FALSE}
#Mapping USA
usa <- map_data("usa")
ggplot() + geom_polygon(data = usa, aes(x=long, y = lat, group = group)) + 
  coord_fixed(1.3)

```

This is the map of the United States. Since we need a map of state, not nationwide, we will get the map of Washington state only.

```{r echo=FALSE}
#Mapping Washington State
states <- map_data("state")
washington <- subset(states, region %in% c("washington"))
ggplot(data = washington) + 
  geom_polygon(aes(x = long, y = lat, group = group), 
               fill = "palegreen", color = "black") + 
  coord_fixed(1.3)

```

This is Washington state! Now we need to look up the counties (subregions).

```{r echo=FALSE}
#divide by county
counties <- map_data("county")
wa_county <- subset(counties, region == "washington")
wa_base <- ggplot(data = wa_county, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + geom_polygon(color = "black", fill = "gray")
wa_base + theme_nothing() + 
  geom_polygon(data = wa_county, fill = NA, color = "white") +
  geom_polygon(color = "black", fill = NA)

```

This seems good, ready to plot on the map. Before doing that, we need to merge the dataset from Google Maps, which has the information of longitude, latitude, and subregion, and the dataset from the Office of Financial Management in Washington, by doing inner join with the common column of subregion. After that, there needs some data cleaning, and get the data of population by simply multiplying density * area.

```{r echo=FALSE}

#mapping
ditch_the_axes <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank()
)

base <- inner_join(wa_county, df, by = "subregion")
no_comma_density <- gsub(",","",base$density)
no_comma_area <- gsub(",","",base$area)
base$density <- as.numeric(no_comma_density)
base$area <- as.numeric(no_comma_area)
base$population <- base$area*base$density

eb1 <- wa_base +
  geom_polygon(data = base, aes(fill = population), color = "white") +
  geom_polygon(color = "black", fill = NA) +
  theme_bw() +
  ditch_the_axes

eb1

```

Here is the map of population. As expected, the county of King has the highest population. We can see the density, too.

```{r echo=FALSE}

eb1_density <- wa_base +
  geom_polygon(data = base, aes(fill = density), color = "white") +
  geom_polygon(color = "black", fill = NA) +
  theme_bw() +
  ditch_the_axes

eb1_density + scale_fill_gradient()

```

As expected, the county of King has the highest population density.

```{r echo=FALSE}

eb2_density <- eb1_density + scale_fill_gradientn(colours = rev(rainbow(7)),
                       breaks = c(2, 4, 10, 50, 100, 1000),
                       trans = "log10")
eb2_density
```

This one looks much better by giving some color.

```{r echo=FALSE}
eb2_density + coord_fixed(xlim = c(-121, -123),  ylim = c(47, 48), ratio = 1.3)
```

Yes, highest population density. Unfortunately, there is no more detail more than a county, we cannot see the area, population, or population density of each districts or zip codes in the county of King.

```{r echo=FALSE, Data_Cleaning_bldg}


bldg = read.csv('Building_Permit.csv')
str(bldg)

bldg$Application.Permit.Number = NULL
bldg$Address = NULL
bldg$Description = NULL
bldg$Applicant.Name = NULL
bldg$Permit.and.Complaint.Status.URL = NULL
bldg$Master.Use.Permit = NULL
bldg$Location = NULL

bldg <- subset(bldg, Category != "")
bldg <- subset(bldg, Action.Type != "")
bldg <- subset(bldg, Status != "")

bldg <- subset(bldg, bldg$Status != "AP Closed" & 
                 bldg$Status != "Permit Closed" &
                 bldg$Status != "CANCELLED")

bldg$Value = as.numeric(gsub("[\\$,]", "", bldg$Value))

sbbox <- make_bbox(lon = bldg$Longitude, lat = bldg$Latitude, f=0.1)
sbbox

sq_map <- get_map(location = sbbox, maptype = "satellite", source = "google")
str(bldg)
```

#Summary of the dataset

Dataset has the list of buildings permitted by the city of Seattle with 55900 rows and 20 variables. For each building, it has its unique application permit number with other characteristics described by variables. It is refreshed daily. More details are available from: https://data.seattle.gov/Permitting/Building-Permits-Current/mags-97de


#Observation

```{r echo=FALSE}
ggmap(sq_map)
```

This is the map of the city of Seattle.

## Permit Type: Is the city of Seattle being developed?

In order to see if the city of Seattle is really being developed, we need to know how many buildings are permitted under construction, demolition, or site development.

```{r echo=FALSE}

p1 <- ggplot(aes(x=Permit.Type), data=bldg) +
  geom_bar()

p2 <- ggmap(sq_map) + 
  geom_point(data = bldg, mapping = aes(x = Longitude, y = Latitude, color=Permit.Type), size=1/2, alpha=1/2) +
  guides(colour = guide_legend(override.aes = list(size=5))) +
    ggtitle("Map by permit type")

grid.arrange(p1, p2, ncol=2)

```

We can see the number of permits under construction is significantly more than demolition, so the city of Seattle is actively being developed as of now.

## Category: How the city of Seattle being developed?

```{r echo=FALSE, Mapping_buildings(univariate)}


p3 <- ggplot(aes(x=Category), data=bldg) +
  geom_bar()

p4 <- ggmap(sq_map) + 
  geom_point(data = bldg, mapping = aes(x = Longitude, y = Latitude, color=Category), size=1/2, alpha=1/2) +
  guides(colour = guide_legend(override.aes = list(size=5))) +
    ggtitle("Map by category")

grid.arrange(p3, p4, ncol=2)

```

There are many building being used as commercial in downtown. we can also see many institutional buildings around the center and north of the city of Washington, where actually University of Washington and Seattle Central College are located. There are two areas that do not have many building in northwest and southwest, because there is a park and a harbor in each. we can also see many industrial building in the south, and that's because there is an industrial district.


## Value: How much are those buildings in the city of Seattle

```{r echo=FALSE}

ggplot(aes(x=Value), data=bldg) +
  geom_bar()

```

This one does not show anything because Value is too inconsistent, needs a transformation by applying log10.

```{r echo=FALSE}

bldg$Value = log10(bldg$Value)
bldg$Value <- cut(bldg$Value, c(0, 2, 3, 4, 5, 6, 8))
bldg_omit_NA <- subset(bldg, Value != "")

p5 <- ggplot(aes(x=Value), data=bldg_omit_NA) +
  geom_bar()

p6 <- ggmap(sq_map) + 
  geom_point(data = bldg_omit_NA, mapping = aes(x = Longitude, y = Latitude, color=Value), size=1/2, alpha=1/2) +
  guides(colour = guide_legend(override.aes = list(size=5))) +
    ggtitle("Map by log10 of value")

grid.arrange(p5, p6, ncol=2)

```

We can see that many buildings in the City of Seattle are valued in between $100,000 - $1,000,000, which is not relatively low nor relatively high compared to other cities. Note that most of expensive buildings are in the center and north of the city of Seattle, where the downtown and so-called "rich area" are located.


# Interpretation

By seeing the map we can see what and how the city of Seattle is being developed.

Yet, it is hard to make sure, because they are mixed up: industrial buildings are obviously more expensive than single house, and the buildings being demolished are surely under-appreciated or have no value at all. The plots following will see the details of each.

## How buildings are being constructed, by category

```{r echo=FALSE, Mapping_buildings_by_properties(multivariate)}

plotMap <- function(k, x)
{
  f <- ggmap(sq_map) + 
    geom_point(data = subset(bldg, bldg$Permit.Type == k & bldg$Category == x), 
                                  mapping = aes(x = Longitude, y = Latitude, color=Action.Type), size=9/10) +
    guides(colour = guide_legend(override.aes = list(size=5))) +
    ggtitle(paste("Correlation between permit type (", k, "), \n category (", x, "), and action type"))
  return(f)
}
                                                                

k <- 'Construction'
c <- 'COMMERCIAL'
ind <- 'INDUSTRIAL'
m <- 'MULTIFAMILY'
s <- 'SINGLE FAMILY / DUPLEX'
ins <- 'INSTITUTIONAL'

pm1 <- plotMap(k, c)

pm2 <- plotMap(k, ind)

pm3 <- plotMap(k, m)

pm4 <- plotMap(k, s)

pm5 <- plotMap(k, ins)

grid.arrange(pm1, pm2, pm3, pm4, pm5, ncol=2)

```

Although many commercial constructions are adding or alternating their building, there are still many new buildings, even in downtown, where the price is highest. Most industrial constructions, on the other hand, are mostly just adding or alternating. For the action type of multifamily, newly constructing buildings are dominating. Constructions of single family are similar to ones of commercial, but the area they have is different. While I do not see many constructions of institutional, I see most of buildings are pointing University of Washington or Seattle Central College.


## How buildings are being demolished, by category


```{r echo=FALSE}
k <- 'Demolition'

pm6 <- plotMap(k, c)

pm7 <- plotMap(k, ind)

pm8 <- plotMap(k, m)

pm9 <- plotMap(k, s)

pm10 <- plotMap(k, ins)

grid.arrange(pm6, pm7, pm8, pm9, pm10, ncol=2)
```

I see that regardless of category of the buildings, there is very few deconstruction or relocation. Also I see that, compared the area that the constructions of multifamily are being built, there are not many demolition of multifamily in downtown. I can predict that there are some constructions of tower blocks, because it is almost impossible to have such high constructions with low demolition, as the area is very limited.

## How buildings are being maintenanced, by category

```{r echo=FALSE}
k <- 'Site Development'

pm11 <- plotMap(k, c)

pm12 <- plotMap(k, ind)

pm13 <- plotMap(k, m)

pm14 <- plotMap(k, s)

pm15 <- plotMap(k, ins)

grid.arrange(pm11, pm12, pm13, pm14, pm15, ncol=2)
```

Most notable thing from these are, there are still many buildings that are being graded.

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}
eb2_density + ggtitle("Density of the state of Washington by county")
```

### Description One

I was surprised how low density some county have. I don't see any particularly large nor small county, but the county of King has more than 1000 density, and some other counties have lower than 4 density. However, I can carefully predict something from here that most of people in the city of Seattle are from other county or state, as its population density imbalance is abnormal.

### Plot Two
```{r echo=FALSE, Plot_Two}
ggplot(aes(x=Permit.Type), data=bldg) +
  geom_bar() +
  ggtitle("Count by permit type")
```

### Description Two

I could somehow predict that the city of Seattle is actively being developed even before I started analysis on it, but didn't know there are that many constructions now. If I see only the number of constructions, and compare it with demolition, I would be safe to say that the city of Seattle is new born city.

### Plot Three
```{r echo=FALSE, Plot_Three}
pm3
```

### Description Three

A lot of multifamily (I assume it is a tower block as its density is extremely high) are adding, alternating, or newly built in downtown and in coastal line. I was expecting, just like downtown in other city, more buildings of commercial and less buildings of multifamily being constructed in downtown.

------

# Reflection

##Issue

First of all, I was able to get and use the map data, there was no data about district, which would contain a lot more detail about the city of Seattle. If there was, I would be able to get a lot better and accurate visualized plots.

Second, I could get a statistics so that could get a sense if, what, and how the city of Seattle is being developed, but nothing beyond it. If I had more sense of analyzing data, I would be able to get something other than just statistics itself, but I was not able to.

At last, although yet I could get meaningful statistics, there are more than half of buildings that are not valid anymore (application closed, permit closed, or just cancelled). I would be able to get better analysis if all of data in the dataset was valid.


##Conclusion

I used three datasets: one about general demography from ggmap, one about specific demography in the state of Washington from the Office of Financial Management in Washington, and last one from the city of Seattle. I explored the dataset and analyzed what category of constructions or demolitions are in progress in which area of the city of Seattle, and what the action type of that construction, destruction, or site development is.

The city of Seattle is actively being developed just like new born city, and it is not limited to only commercial, but also residential, institutional, and other purposes. Yet most of buildings, except in the area of downtown, are not as expensive as New York City or San Francisco Bay area.

If there was enough time, I would also compare those buildings by yearly, so can find out if this kind of construction boom is recent, or historically has been done like this.