---
output:
  html_document: default
  pdf_document: default
---
The Quality of Wines and its Characteristics by Nicholas Huh
========================================================


```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

suppressMessages(library(ggplot2))
suppressMessages(library(gridExtra))
suppressMessages(library(corrplot))
suppressMessages(library(dplyr))
```

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r echo=FALSE, Load_the_Data}

# Load the Data
setwd('C:/Users/admin/Desktop/Nanodegree/R/Project 5')
ds <- read.csv('wineQualityReds.csv')
ds$X = NULL
```

This report explores a dataset containg acid and ingredients, % of alcohol, and the quality of wine with 1599 rows.


# Univariate Plots Section

```{r echo=FALSE, Univariate_Plots}

# knitr::opts_chunk$set(fig.width=9, fig.height=6, echo=FALSE, warning=FALSE, message=FALSE)

names(ds)

str(ds)

summary(ds)

dim(ds)

```

My dataset consists of 1599 rows with 11 variables. In order to see the statistics and distibution of each variable, I will crate histograms for each variables, in total of 12, as well as summaries.

```{r echo=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
#Properties

p1 <- ggplot(aes(x=quality), data=ds) + geom_bar() + ggtitle('Quality of wines') + 
  theme(plot.title=element_text(hjust = 0.5))

p2 <- ggplot(aes(x=pH), data=ds) + geom_bar(binwidth=0.1) + 
  scale_x_continuous(breaks = c(2.8, 3.0, 3.2, 3.4, 3.6, 3.8, 4.0)) +
  ggtitle('Level of pH') + theme(plot.title=element_text(hjust = 0.5))

p3 <- ggplot(aes(x=alcohol), data=ds) + geom_bar(binwidth=0.1) + 
  scale_x_continuous(breaks = c(8, 8.5, 9, 9.5, 10, 11, 12, 13, 14)) +
  ggtitle('% of alcohol') + theme(plot.title=element_text(hjust = 0.5))

p4 <- ggplot(aes(x=density), data=ds) + geom_bar(binwidth=0.001) + 
  ggtitle('Level of density') + 
  theme(plot.title=element_text(hjust = 0.5))

grid.arrange(p1, p2, p3, p4, ncol=2)

summary(ds$pH)

summary(ds$alcohol)

summary(ds$quality)

```

I can see most of wines have 5-6 quality in the dataset. Also I can see that the lowest quality is 3, highest quality is 8, median is 6, mean is 5.636 Likewise above, I can see pH sits in between 3.0 and 3.6 with lowest pH is 2.74, highest pH is 4.010, mean is 3.311, median is 3.310. Most of wines have 9.0-12.0 % of alcohol with lowest of 8.40, highest of 14.90, mean of 10.42, and median of 10.20. I am not sure how density can relate to the quality of wines, but most wines sit on between 0.995 and 1.000. Next I will see all the kinds of acid.

```{r echo=FALSE}
#Acidity 

p5 <- ggplot(aes(x=fixed.acidity), data=ds) + geom_bar(binwidth=0.2) + 
  scale_x_continuous(breaks = c(4, 6, 7, 8, 9, 10, 12, 14, 16)) + 
  ggtitle('Level of fixed acidity') + 
  theme(plot.title=element_text(hjust = 0.5))

p6 <- ggplot(aes(x=volatile.acidity), data=ds) + 
  geom_bar(binwidth=0.02) + 
  scale_x_continuous(breaks = c(0.2, 0.4, 0.6, 0.8, 1.0, 1.2, 1.6)) + 
  ggtitle('Level of volatile acidity') + 
  theme(plot.title=element_text(hjust = 0.5))

p7 <- ggplot(aes(x=citric.acid), data=ds) + 
  geom_bar(binwidth=0.03) + 
  scale_x_continuous(breaks = c(0.0, 0.2, 0.4, 0.6, 0.8, 1.0)) + 
  ggtitle('Level of citric acid') + 
  theme(plot.title=element_text(hjust = 0.5))

grid.arrange(p5, p6, p7, ncol=2)

```

I can see most of wines have 6-10 g/dm^3 of fixed acidity and 0.2-0.8 g/dm^3 of volatile acidity. Citric acid looks a little bit different as there are a lot of wines that have no citric acid at all. Including that, most wines have 0.0-0.6 g/dm^3 of citric acid.

The following plots of each ingredient, however, need transformation for better understanding.

```{r echo=FALSE}

#Ingredients

p8 <- ggplot(aes(x=residual.sugar), data=ds) + 
  geom_bar(binwidth=0.4) + 
  scale_x_continuous(breaks = c(1, 2, 3, 4, 8, 12, 16)) + 
  ggtitle('Amount of residual sugar') + 
  theme(plot.title=element_text(hjust = 0.5))

p9 <- ggplot(aes(x=residual.sugar), data=ds) + 
  geom_bar(binwidth=0.1) + 
  scale_x_log10(breaks=c(1, 1.5, 2, 3, 4, 5)) + 
  ggtitle('Log10 of amount of residual sugar') + 
  theme(plot.title=element_text(hjust = 0.5))

p10 <- ggplot(aes(x=chlorides), data=ds) + 
  geom_bar(binwidth=0.01) + 
  scale_x_continuous(breaks = c(0.0, 0.05, 0.1, 0.15, 0.2, 0.4, 0.6)) + 
  ggtitle('Amount of chlorides') + 
  theme(plot.title=element_text(hjust = 0.5))

p11 <- ggplot(aes(x=chlorides), data=ds) + 
  geom_bar(binwidth=0.02) + 
  scale_x_log10(breaks=c(0.01, 0.05, 0.08, 0.1, 0.2)) + 
  ggtitle('Log10 of amount of chlorides') + 
  theme(plot.title=element_text(hjust = 0.5))

p12 <- ggplot(aes(x=free.sulfur.dioxide), data=ds) + 
  geom_bar(binwidth=2) + 
  scale_x_continuous(breaks = c(0, 5, 10, 15, 20, 25, 30, 35, 40, 60)) + 
  ggtitle('Amount of free sulfur dioxider') + 
  theme(plot.title=element_text(hjust = 0.5))

p13 <- ggplot(aes(x=free.sulfur.dioxide), data=ds) + 
  geom_bar(binwidth=0.08) + 
  scale_x_log10(breaks=c(1, 5, 7, 10, 15, 20, 30)) + 
  ggtitle('Log10 of free sulfur dioxide') + 
  theme(plot.title=element_text(hjust = 0.5))

p14 <- ggplot(aes(x=total.sulfur.dioxide), data=ds) + 
  geom_bar(binwidth=8) + 
  ggtitle('Amount of total sulfur dioxide') + 
  scale_x_continuous(breaks = c(20, 40, 60, 80, 100, 120, 140, 160, 180, 200, 300)) + 
  theme(plot.title=element_text(hjust = 0.5))

p15 <- ggplot(aes(x=total.sulfur.dioxide), data=ds) + 
  geom_bar(binwidth=0.08) + 
  scale_x_log10(breaks=c(5, 10, 20, 25, 40, 55, 70, 100, 120, 150)) + 
  ggtitle('Log10 of amount of total sulfur dioxide') + 
  theme(plot.title=element_text(hjust = 0.5))

p16 <- ggplot(aes(x=sulphates), data=ds) + 
  geom_bar(binwidth=0.04) + 
  scale_x_continuous(breaks = c(0.2, 0.4, 0.6, 0.8, 1, 1.5, 2.0)) + 
  ggtitle('Amount of sulphates') + 
  theme(plot.title=element_text(hjust = 0.5))

p17 <- ggplot(aes(x=sulphates), data=ds) + 
  geom_bar(binwidth=0.02) + 
  scale_x_log10(breaks=c(0.2, 0.4, 0.5, 0.6, 0.7, 0.8, 1, 1.2, 1.5)) + 
  ggtitle('Log10 of amount of sulphates') + 
  theme(plot.title=element_text(hjust = 0.5))

grid.arrange(p8, p9, p10, p11, p12, p13, p14, p15, p16, p17, ncol=2)

```

It is negatively skewed and hard to understand, try log10 and be able to get much better result. I can see it has distribution between 1.5-3.0 g/dm^3 the most. Similar to the case above, I apply transformation and can see most wines have 0.05-0.10 g/dm^3 of chlorides. I apply transformation and can see most wines have 0.05-0.10 g/dm^3 of free sulfur dioxide. By applying transformation I can see most wines have under 150 mg/dm^3 of total sulfur dioxide. Most wines have 0.5-0.8 g/dm^3 of sulphates.


# Univariate Analysis

### What is the structure of your dataset?

There are 1,599 red wines with 11 variables (3 kinds of acid, 5 of ingredients, density, pH, % of alcohol, and quality).

Other observations: \
  * Most wines are of 5 or 6 quality. \
  * The median % of alcohol is 10.20. \
  * Most wines sit on the pH of 3.0-3.5. \
  * All ingredients shows positively skewed graph. \
  * None of variable has uniform distribution. \



### What is/are the main feature(s) of interest in your dataset?

The most important feature in the dataset is the quality of wine. I'd like to define what factors affect the quality of wine, the most and the least.


### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

All the variables possibly affect the quality of wine, especially ingredients as they affect the taste of wines.

### Did you create any new variables from existing variables in the dataset?

I didn't create new variable as they are well-organized and there is no specific and well-defined correlation yet.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

I performed "scale_x_log10()", which transforms data by applying log10. By doing so, I can see the distribution more clearly as the shape of graph transforms from abnormal form (such as negatively skewed or positively skewed) to more likely standard destribution. I've applied it to all of ingredients.



# Bivariate Plots Section

First thing to do is to create a map of correlation, so that I can understand how each variables are related to other variables, including the quality of wines which I need to find out.

```{r echo=FALSE, Bivariate_Plots}

M <- cor(ds)
corrplot(M, method="circle")

```

I can see my correlation, and I can predict that % of alcohol and the level of volatile acidicty affect the most among them.

I will use jitter plot for the first 4 plots of alcohol, residual sugar, chlorides, total sulfur dioxide, free sulfur dioxide, and sulphates, which are not ingredients as they didn't need transformation.

```{r echo=FALSE}

pc1 <- ggplot(ds, aes(x = quality, y = alcohol)) + 
  geom_jitter() + 
  scale_x_continuous(breaks = c(3, 4, 5, 6, 7, 8)) + 
  ggtitle('Correlation between quality and alcohol') + 
  theme(plot.title=element_text(hjust = 0.5))

pc2 <- ggplot(ds, aes(x = quality, y = residual.sugar)) + 
  geom_jitter() + 
  scale_x_continuous(breaks = c(3, 4, 5, 6, 7, 8)) +
  ggtitle('Correlation between quality and residual sugar') + 
  theme(plot.title=element_text(hjust = 0.5))

pc3 <- ggplot(ds, aes(x = quality, y = residual.sugar)) + 
  geom_jitter(alpha=1/10) + 
  scale_x_continuous(breaks = c(3, 4, 5, 6, 7, 8)) +
  ggtitle('Correlation between quality and log10 of residual sugar') + 
  theme(plot.title=element_text(hjust = 0.5))

pc4 <- ggplot(ds, aes(x = quality, y = chlorides)) + 
  geom_jitter() + 
  scale_x_continuous(breaks = c(3, 4, 5, 6, 7, 8)) +
  ggtitle('Correlation between quality and chlorides') + 
  theme(plot.title=element_text(hjust = 0.5))

pc5 <- ggplot(ds, aes(x = quality, y = total.sulfur.dioxide)) + 
  geom_jitter() + 
  scale_x_continuous(breaks = c(3, 4, 5, 6, 7, 8)) +
  ggtitle('Correlation between quality and total sulfur dioxide') + 
  theme(plot.title=element_text(hjust = 0.5))

pc6 <- ggplot(ds, aes(x = quality, y = sulphates)) + 
  geom_jitter() + 
  scale_x_continuous(breaks = c(3, 4, 5, 6, 7, 8)) +
  ggtitle('Correlation between quality and sulphates') + 
  theme(plot.title=element_text(hjust = 0.5))

pc7 <- ggplot(ds, aes(x = quality, y = free.sulfur.dioxide)) + 
  geom_jitter() + 
  scale_x_continuous(breaks = c(3, 4, 5, 6, 7, 8)) +
  ggtitle('Correlation between quality and free sulfur dioxide') + 
  theme(plot.title=element_text(hjust = 0.5))

grid.arrange(pc1, pc2, pc3, pc4, pc5, pc6, pc7, ncol=2)

```

I can see that higher quality of wines tend to have higher alcohol %. Most of wines have 2-3 g/dm^3, and the amount of sugar doesn't affact to the quality of wine much. However, this one is a bit arguable, so will explore more on this specific variable by transforming the data. Even after transforming the data, I still can't see significant correlation. Most of wines have 0.05-0.1 g/dm^3, and the amount of chlorides doesn't affact to the quality of wine much. The amount of sulfur dioxide is vary, and higher quality of wines tend to contain less amount of sulfur dioxide. Most of wines have around 0.5-1 g/dm^3, and higher quality of wines tend to contain more amount of sulphates. I can't see any significant correlation between quality and free sulfur dioxide.

From this point, I will try to use box plot instead of point plot, as they were ingredients that needed transformation and box plot shows the median and quantiles.

```{r echo=FALSE}
pb1 <- ggplot(ds, aes(x = quality, y = fixed.acidity)) + 
  geom_boxplot(aes(group=quality)) + 
  scale_x_continuous(breaks = c(3, 4, 5, 6, 7, 8)) +
  ggtitle('Correlation between quality and fixed acidity') + 
  theme(plot.title=element_text(hjust = 0.5))

pb2 <- ggplot(ds, aes(x = quality, y = volatile.acidity)) + 
  geom_boxplot(aes(group=quality)) + 
  scale_x_continuous(breaks = c(3, 4, 5, 6, 7, 8)) +
  ggtitle('Correlation between quality and volatile acidity') + 
  theme(plot.title=element_text(hjust = 0.5))

pb3 <- ggplot(ds, aes(x = quality, y = citric.acid)) + 
  geom_boxplot(aes(group=quality)) + 
  scale_x_continuous(breaks = c(3, 4, 5, 6, 7, 8)) +
  ggtitle('Correlation between quality and citric acid') + 
  theme(plot.title=element_text(hjust = 0.5))

pb4 <- ggplot(ds, aes(x = quality, y = pH)) + 
  geom_boxplot(aes(group=quality)) + 
  scale_x_continuous(breaks = c(3, 4, 5, 6, 7, 8)) +
  ggtitle('Correlation between quality and pH') + 
  theme(plot.title=element_text(hjust = 0.5))

pb5 <- ggplot(ds, aes(x = quality, y = density)) + 
  geom_boxplot(aes(group=quality)) + 
  scale_x_continuous(breaks = c(3, 4, 5, 6, 7, 8)) +
  ggtitle('Correlation between quality and density') + 
  theme(plot.title=element_text(hjust = 0.5))

grid.arrange(pb1, pb2, pb3, pb4, pb5, ncol=2)
```

I don't see significant correlation between quality and fixed acidity. Higher quality of wines tends to have lower level of volatile acidity. Higher quality of wines tends to have higher level of volatile acidity. Higher quality of wines tends to have lower level of pH. Higher quality of wines tends to have slightly lower level of density.

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

The most important variable of dataset is the quality of wine, thus all the correlation above is between the quality and other variables. With the analysis above, I could get the following facts: \

  The quality of wines has strong positive proportional with:\
    * alcohol %\
    * the amount of sulphates\
    * the level of citric acid\
  Weak positive proportional with:\
    * the level of fixed acidity\
  No significant relationship with:\
    * the amount of residual sugar\
    * the level of free sulfur dioxide\
    * the level of pH\
  Weak negative proportinal with:\
    * the amount of chlorides\
    * the amount of total surfur dioxide\
    * the level of density\
  Strong negative proportinal with:\
    * the level of volatile acidity\
    

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

Relationship between the level of pH and the level of fixed acidity, which has as much correlation as -0.8, has the most correlation among all the correlations.

### What was the strongest relationship you found?

The quality of wine has the strongest relationship with the level of volatile acidity and the level of alcohol. It also has some relationship with the amount of sulphates and the level of citric acid.


# Multivariate Plots Section

From the data above, I can tentatively conclude that % of alcohol, the amount of sulphates, and level of citric acid and volatile acidity are the biggest factors on the quality of wines. The following plots will prove if they are true or not.

```{r echo=FALSE, Multivariate_Plots}

ds$alcohol.bucket <- cut(ds$alcohol, c(8, 9, 10.5, 12, 16))

ggplot(aes(x = factor(quality), y = volatile.acidity), data = ds) + 
   geom_boxplot(aes(fill = alcohol.bucket)) + 
  ggtitle('Correlation between alcohol, quality, and volatile acidity') + 
  theme(plot.title=element_text(hjust = 0.5)) + 
  scale_x_discrete("Quality")
  
```

First I cut alcohol to make it a factor, then create a boxplot that accepts three variables: alcohol, quality, volatile acidity, which are two most important factors according to correlation. I can see that higher quality of wines have higher alcohol (which is purple) and lower volatile acidity.

```{r echo=FALSE}
p1 <- ggplot(aes(x = alcohol, fill = as.factor(quality), group=quality), 
             data = ds) +
  geom_density(adjust=1.5, position='fill') +
  scale_fill_discrete(name = "Quality") +
  scale_fill_brewer()

p2 <- ggplot(aes(x = volatile.acidity, fill = as.factor(quality), 
                 group=quality), data = ds) +
  geom_density(adjust=1.5, position='fill') +
  scale_fill_discrete(name = "Quality") +
  scale_fill_brewer()

p3 <- ggplot(aes(x = sulphates, fill = as.factor(quality), group=quality), 
             data = ds) +
  geom_density(adjust=1.5, position='fill') +
  scale_fill_discrete(name = "Quality") +
  scale_fill_brewer()

p4 <- ggplot(aes(x = citric.acid, fill = as.factor(quality), group=quality),
             data = ds) +
  geom_density(adjust=1.5, position='fill') +
  scale_fill_discrete(name = "Quality") +
  scale_fill_brewer()

grid.arrange(p1, p2, p3, p4, ncol=2, top="Multivariate plots by density")
```

Plot above is a summerise of 4 top correlation variables (alcohol, volatile acidity, sulhpates, and citric acid.) Now everything is clear.

```{r echo=FALSE}

ggplot(ds, aes(fill=factor(quality), y=volatile.acidity, x=alcohol.bucket)) +
  geom_bar(stat="identity", position="fill") +
  scale_fill_brewer()

```

I also created stack barplot, by the percentage. I can see that most of quality of 8 sits on 12-16% of alcohol and under 0.25 of volatile acidity.

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

With the analysis above, I can conclude that, in addition to the proportion of each variables, at which point the quality of wines peaks.

### Were there any interesting or surprising interactions between features?

I was surprised that even if volatile acidity has the strongest negative proportional with the quality of wines, still there needs some (approximately 0.2 g/dm^3)

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

I didn't create any models.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}
ggplot(aes(x=citric.acid), data=ds) + geom_histogram(binwidth=0.02) + 
  ggtitle('Level of citric acid') + theme(plot.title=element_text(hjust = 0.5)) +
  scale_x_continuous("Citric Acid [g / dm^3]")
  
```

### Description One

There are so many wines that have 0 citric acid. It is still curious if that's because they have no data or actually there are wines that have no citric acid at all.

### Plot Two
```{r echo=FALSE, Plot_Two}

ggplot(ds, aes(x = quality, y = alcohol)) + geom_jitter(ylab='Alcohol [%]') + 
  scale_x_continuous(breaks = c(3, 4, 5, 6, 7, 8)) + 
  ggtitle('Correlation between quality and alcohol') + 
  theme(plot.title=element_text(hjust = 0.5)) + 
  scale_y_continuous("Alcohol [%]")

```

### Description Two

It is interesting how alcohol affects the quality of wines. The percentage of alcohol is normally, unless it is abnormally high or low, not considered as one of factors that decides the quality of wines, but actually it is the most important factor among all other chemical components.

### Plot Three
```{r echo=FALSE, Plot_Three}
ggplot(aes(x = volatile.acidity, fill = as.factor(quality), group=quality), 
       data = ds) +
  geom_density(adjust=1.5, position='fill') + 
  ggtitle('Density by quality and volatile acidity') + 
  theme(plot.title=element_text(hjust = 0.5)) + 
  scale_fill_discrete(name = "Quality") +
  scale_x_discrete("Volatile Acidity [g / dm^3]") +
  scale_fill_brewer()
```

### Description Three

As described on multivariate plots, it is surprising that even though volatile acidity is a chemical component that can lead to an unpleasant, vinegar taste, it is still necessary for the good quality of wines.\

------

# Reflection

This report explores a dataset containg acid and ingredients, % of alcohol, and the quality of wine with 1599 rows. I explored the dataset in order to find out which factors affect positively or negatively, and which one does the most and the least.

I had struggle when I try to find more than 3 relationships, because all the variables seem independent. However yet I was successfully found what decides the quality of wines and how much it actually affects.

After exploring dataset, the correlation between the quality of wines and other factors are clear, being able to predict what the quality of a certain wine will be by looking up those factors. The most interesting fact is, if the percent of alcohol of a wine is lower than average, it is very likely rated as low quality.

It was able to get a good result from this dataset by using R, but I could get the better results if I had more data, as it had only 1599 rows. I would be able to get the better result if there was a variable about what the wine was made of, such as the type and origin of grapes and the county of manufacture.